\sysfull

### 2019-06-22 paper/TOSN/evaluation.tex
TOSN coala version
minor revision TOSN

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/coalEfficiency}
    \caption{Breakdown of overhead for the proposed coalescing strategies:
    NC (no coalescing), EO (energy-oblivious), EG (energy-guided), WEG
    (weighted energy-guided). All coalescing strategies reduce total
    overhead and maximize useful work. EG and WEG perform better than EO.}
    \label{fig:overallOverheadBreakdown}
\end{figure}

@@

\begin{figure}
    \begin{subfigure}{\columnwidth}
        \includegraphics[width=\columnwidth]{figures/overallOverhead.pdf}
        \caption{Kernel overhead breakdown}
        \label{fig:kernel-overhead}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \includegraphics[width=\columnwidth]{figures/memAccess.pdf}
        \caption{Protected memory accesses breakdown}
        \label{fig:mem-accesses}
    \end{subfigure}
    \caption{\sys's internal overhead. NC: no coalescing, EG:
    energy-guided, WEG: weighted energy-guided.}
    \label{fig:war}
\end{figure}

@@

\paragraph{Speedup with Coalescing.}

Fig.~\ref{fig:coalescing} shows \sys's run time of two coalescing
strategies (EG, WEG) normalized to the run time without coalescing (NC).
%
The results show that all benchmarks
complete faster with coalescing than without coalescing: from 25\%
(\textit{ar}) up to 70\% (\textit{sort}).
%
This speedup is a consequence of the reduced overhead demonstrated in
Section~\ref{sec:coala_overhead}.
%
However, the magnitude of the speedup is (1) highly application-dependent and
(2) largely similar across the two coalescing strategies, with the exception of
\textit{fft}.
%
In some cases (\textit{bc}, \textit{cuckoo}, \textit{sort}) WEG's task weighting
system is counter-productive.
%
This occurs in task decompositions with energy-uniform tasks, where counting
tasks disregarding their energy consumption provides an equal amount of information
with a smaller effort.
%
In \textit{fft}, tasks are not uniform, and accounting for their different weights
beneficial. In fact, the lack of task energy awareness is detrimental: with EG
\textit{fft} runs slower than without any coalescing (NC).
%
The speedup is highest for \textit{bc}, \textit{cuckoo}, \textit{dijkstra} and
\textit{sort}, because their tasks are relatively small and are easily coalesced.
%
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/coalStrategies}%
    \caption{\sys's coalescing performance. Application execution time
with coalescing (EG and WEG) normalized to the execution time
without coalescing (NC).}
\label{fig:coalescing}
\end{figure}

\begin{figure}
    \includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
    \caption{\sys's execution time normalized to Alpaca's,
    for three distances from the energy source (15, 30 and 50\,cm).}
    \label{fig:runtime}
\end{figure}

\paragraph{Benefits of Adaptive Tasks.}
%
We now compare \sys's performance to Alpaca~\cite{alpaca}---a
\emph{non-adaptive} task-based system with tasks fixed at compile-time.
Fig.~\ref{fig:runtime} shows the average execution time of each
application for \sys and Alpaca, normalized, where possible, to the latter.
\sys provides a performance benefit compared to Alpaca for most
applications. The speedup is greatest for applications with repeated WAR
dependencies throughout their code, particularly involving arrays
(\textit{dijkstra}, \textit{fft} and \textit{sort}). \sys's VMM
successfully amortizes the overhead of protecting memory that is accessed in
such patterns.  In applications without locality among accesses to protected
variables, including array elements, \sys incurs overhead from memory
virtualization that causes its performance to be comparable to (or worse than)
Alpaca (\textit{ar}, \textit{cuckoo}).

Due to its static progressing behavior, Alpaca was unable to complete the \textit{fft} benchmark on distances larger than 15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
execution is negligible and stored energy is insufficient to complete some of
the static tasks.}. This is marked with $\infty$ signs in Fig.~\ref{fig:coalescing}. \sys, however, manged to complete \textit{fft} by enabling its task downscaling at 30\,cm and 50\,cm from the energy source.

@@

\begin{figure}
    \begin{subfigure}{\columnwidth}
        \includegraphics[width=\columnwidth]{figures/page_exec-time.pdf}
        \caption{Execution time normalized to lowest per-application}
        \label{fig:page-exec-time}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \includegraphics[width=\columnwidth]{figures/pagePulls.pdf}
        \caption{Number of page faults per application run}
        \label{fig:page-pulls}
    \end{subfigure}
    \caption{Effect of page size (in bytes).}
    \label{fig:war}
\end{figure}

>>>>

\begin{figure}
    \centering
    \includegraphics[width=.8\columnwidth]{figures/coalEfficiency}
    \caption{Breakdown of overhead for the proposed coalescing strategies:
    NC (no coalescing), EO (energy-oblivious), EG (energy-guided), WEG
    (weighted energy-guided). All coalescing strategies reduce total
    overhead and maximize useful work. EG and WEG perform better than EO.}
    \label{fig:overallOverheadBreakdown}
\end{figure}

@@

\begin{figure}
    \begin{subfigure}{\columnwidth}
			\centering
        \includegraphics[width=.8\columnwidth]{figures/overallOverhead.pdf}
        \caption{Kernel overhead breakdown}
        \label{fig:kernel-overhead}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
			\centering
        \includegraphics[width=.8\columnwidth]{figures/memAccess}
        \caption{Protected memory accesses breakdown}
        \label{fig:mem-accesses}
    \end{subfigure}
    \caption{\sys's internal overhead. NC: no coalescing, EG:
    energy-guided, WEG: weighted energy-guided.}
    \label{fig:war}
\end{figure}

@@

\paragraph{Speedup with Coalescing.}
%
\begin{figure}
    \includegraphics[width=.8\columnwidth]{figures/coalStrategies}%
    \caption{\sys's coalescing performance. Application execution time
with coalescing (EG and WEG) normalized to the execution time
without coalescing (NC).}
\label{fig:coalescing}
\end{figure}

\begin{table}[t]%

\caption{\add{Comparison of \sys performance on two different capacitor sizes. We see that \sys optimizes its coalesced task size based on the energy buffer size.
\textbf{coalesced task} refers to the first coalesced task after a reboot, \textbf{exp. time} shows the experiment duration, and the \textbf{runs} column lists 
the number of complete runs of the \textit{sort} application during the experiments.}
}
\label{tab:dif_cap}
\add{
\begin{minipage}{\columnwidth}
\begin{center}
\begin{tabular}{llllll}
  \toprule
    \textbf{cap. size} ($\mu\text{f}$)& \textbf{exp. time} ($s$)& \textbf{duty cycle} & \textbf{coalesced task} ($ms$)   & \textbf{runs} & \textbf{app run-time} ($ms$) \\
        \hline
    47                 &1205             & 12.93\%    & 33                        & 85       & 183      \\
    470                &1205             & 12.79\%    & 88                        & 87       & 177 \\
  \bottomrule
\end{tabular}
\end{center}
\end{minipage}  
}
\end{table}%

Fig.~\ref{fig:coalescing} shows \sys's run time of two coalescing
strategies (EG, WEG) normalized to the run time without coalescing (NC).
%
The results show that all benchmarks
complete faster with coalescing than without coalescing: from 25\%
(\textit{ar}) up to 70\% (\textit{sort}).
%
This speedup is a consequence of the reduced overhead demonstrated in
Section~\ref{sec:coala_overhead}.
%
However, the magnitude of the speedup is (1) highly application-dependent and
(2) largely similar across the two coalescing strategies, with the exception of
\textit{fft}.
%
In some cases (\textit{bc}, \textit{cuckoo}, \textit{sort}) WEG's task weighting
system is counter-productive.
%
This occurs in task decompositions with energy-uniform tasks, where counting
tasks disregarding their energy consumption provides an equal amount of information
with a smaller effort.
%
In \textit{fft}, tasks are not uniform, and accounting for their different weights
beneficial. In fact, the lack of task energy awareness is detrimental: with EG
\textit{fft} runs slower than without any coalescing (NC).
%
The speedup is highest for \textit{bc}, \textit{cuckoo}, \textit{dijkstra} and
\textit{sort}, because their tasks are relatively small and are easily coalesced.
%

\begin{figure}
    \includegraphics[width=.8\columnwidth]{figures/coala_alpaca_gcc}
    \caption{\sys's execution time normalized to Alpaca's,
    for three distances from the energy source (15, 30 and 50\,cm). }
    \label{fig:runtime}
\end{figure}

\paragraph{\add{Different Capacitor Sizes}}
\add{
Table~\ref{tab:dif_cap} shows how \sys optimizes its Coalescing task size based on the amount of buffered energy.
We see that \sys scales up its coalesced task size with a bigger energy buffer and vise versa. This allows it to reduce the time-to-completion of the applications. For example, 
the \textit{sort} run-time is reduced from 183\,$ms$ to 177\,$ms$ when the capacitor is changed from 47\,$\mu$f to 470\,$\mu$f. It should be emphasized that a device with a bigger energy buffer suffers less from power failures (but requires longer charging time). However, as has been measured in Fig.~\ref{fig:overallOverheadBreakdown} and explained in Fig.~\ref{fig:coal_anatomy}, the re-execution penalty for the \textit{sort} application is negligible.  
}

\paragraph{Benefits of Adaptive Tasks.}
%
We now compare \sys's performance to Alpaca~\cite{alpaca}---a
\emph{non-adaptive} task-based system with tasks fixed at compile-time.
Fig.~\ref{fig:runtime} shows the average execution time of each
application for \sys and Alpaca, \add{normalized to the latter or, when not possible, to one second}.
\sys provides a performance benefit compared to Alpaca for most
applications. \add{For example, it is 54\% faster than Alpaca when executing the \textit{sort} application. In general,}  the speedup is greatest for applications with repeated WAR
dependencies throughout their code, particularly involving arrays
(\textit{dijkstra}, \textit{fft} and \textit{sort}). \sys's VMM
successfully amortizes the overhead of protecting memory that is accessed in
such patterns.  In applications without locality among accesses to protected
variables, including array elements, \sys incurs overhead from memory
virtualization that causes its performance to be comparable to (or worse than)
Alpaca (\textit{ar}, \textit{cuckoo}).

Due to its static progressing behavior, Alpaca was unable to complete the \textit{fft} benchmark on distances larger than 15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
execution is negligible and stored energy is insufficient to complete some of
the static tasks.}. This is marked with $\infty$ signs in Fig.~\ref{fig:runtime}. \sys, however, managed to complete \textit{fft} by enabling its task downscaling at 30\,cm and 50\,cm from the energy source.

@@

\begin{figure}
	\centering
    \begin{subfigure}{\columnwidth}
			\centering
        \includegraphics[width=.8\columnwidth]{figures/page_exec-time.pdf}
        \caption{Execution time normalized to lowest per-application}
        \label{fig:page-exec-time}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
			\centering
        \includegraphics[width=.8\columnwidth]{figures/pagePulls.pdf}
        \caption{Number of page faults per application run}
        \label{fig:page-pulls}
    \end{subfigure}
    \caption{Effect of page size (in bytes).}
    \label{fig:war}
\end{figure}

### 2019-06-22 paper/TOSN/hwswsetup.tex
TOSN coala version
minor revision TOSN

\subsection{Experimental Setup}
\label{sec:results_hardware_software}

We used two platforms to evaluate \sys: (i) an \emph{intermittently-powered}
energy-harvesting device (WISP\,5.1~\cite{wisp5,wisp}), and (ii) a \emph{continuously-powered} evaluation
board (MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad}).
%
Every benchmark used in Section~\ref{sec:evaluation} was run repeatedly on
each platform for a few minutes to ensure capturing a diverse power 
trace and produce a high number of complete runs of all applications.
The results were averaged across the duration of the experiment.

WISP contains the MSP430FR5969~\cite{wolverine} MCU with 64\,kB of
non-volatile (FRAM) memory and 2\,kB of volatile (SRAM) memory and was
configured to 1\,MHz clock speed.
%
We powered WISP using an RF signal generator emitting a 20\,dBm sinusoidal wave at 915\,MHz.
The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
antenna~\cite{atlas2015}.
%
The antenna was oriented towards and in parallel with WISP's antenna, and
no objects obstructed the path.
%
We affixed WISP with a paper harness at the edge of a table at a height, from the table suface, of 10\,cm.
%
For distance-controlled experiments we positioned WISP at $d=\{15, 30,
50\}$\,cm from the exciter antenna.
%
% We chose relatively shorter distances than in a deployment scenario to reduce
% charge time and run experiments faster, after having confirmed ourselves a
% result also reported in prior work~\cite{alpaca} that the amount of energy
% available in one charge cycle---the sum of stored energy and incoming energy
% during execution---is the same at distances beyond 50\,cm where incoming energy
% becomes negligible.
%
To obtain execution time, the software toggled GPIO pins at sections of the code
under profile, and the Saleae~\cite{saleae} logic analyzer measured
intervals between edges in the signal. For continous power experiments, the 
execution time was measured using the clock features in TI Code Composer Studio IDE version 7.1.
%
\subsection{Software Benchmarks}
\label{sec:software_benchmarks}

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{| r|r|r | p{0.54\columnwidth} |}
		\hline
		App.&Tasks&SLOC&Description\\
		\hline\hline
        \emph{ar} &10 &428 & activity recognition using a KNN\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
		\hline
        \emph{bc} &10 &371 & several bitcount algorithms\\
        % \textbf{bc} &10 &371 & Counts bits in bytes sequences
% using several algorithms and check for correctness\\
		\hline
        \emph{cuckoo} &14 &426 & Cuckoo Filter with pseudo-random values\\
		\hline
        \emph{dijkstra} &5 &198 & Dijkstra shortest path algorithm \\
		\hline
        \emph{fft} &8 &449 & Fast Fourier Transform\\ % on three pre-generated 128 sample vectors\\
		\hline
		\emph{sort} &4 &167 & selection sort algorithm\\
		\hline
	\end{tabular}
\caption{Characteristics of benchmarks used for evaluation.}
\label{table:benchmark_table}
\end{table}

>>>>

\subsection{Experimental Setup}
\label{sec:results_hardware_software}
\add{
We used three different setups to evaluate \sys: 
(i) an RF-powered energy-harvesting device, WISP\,5.1~\cite{wisp5,wisp}, with a fixed capacitor size of 47\,$\mu$f; 
(ii) an MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad} powered by a BQ25570~\cite{BQ25570EVM-206_website} solar power harvester that is connected to an IXYS SLMD121H04L solar cell~\cite{SLMD121H04L_website} for experimenting with different energy buffers; and
(iii) an MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad} that is continuously powered.
}

\add{
Every benchmark used in Section~\ref{sec:evaluation} was run repeatedly on
each platform for a few minutes to ensure capturing a diverse power 
trace. The exact number of iterations depends on the ambient power intensity and the application itself.
In our experiments, the number of complete runs ranges from 4 to 125.
}

WISP contains the MSP430FR5969~\cite{wolverine} MCU with 64\,kB of
non-volatile (FRAM) memory and 2\,kB of volatile (SRAM) memory and was
configured to 1\,MHz clock speed.
%
We powered WISP using an RF signal generator emitting a 20\,dBm sinusoidal wave at 915\,MHz.
The signal generator was connected to the Laird RFMAX S9028PCRJ 8\,dBic
antenna~\cite{atlas2015}.
%
The antenna was oriented towards and in parallel with WISP's antenna, and
no objects obstructed the path.
%
We affixed WISP with a paper harness at the edge of a table at a height, from the table surface, of 10\,cm.
%
For distance-controlled experiments we positioned WISP at $d=\{15, 30,
50\}$\,cm from the exciter antenna.
%
% We chose relatively shorter distances than in a deployment scenario to reduce
% charge time and run experiments faster, after having confirmed ourselves a
% result also reported in prior work~\cite{alpaca} that the amount of energy
% available in one charge cycle---the sum of stored energy and incoming energy
% during execution---is the same at distances beyond 50\,cm where incoming energy
% becomes negligible.
%
To obtain execution time, the software toggled GPIO pins at sections of the code
under profile, and the Saleae~\cite{saleae} logic analyzer measured
intervals between edges in the signal. For continuous power experiments, the 
execution time was measured using the clock features in TI Code Composer Studio IDE version 7.1.
%
\subsection{Software Benchmarks}
\label{sec:software_benchmarks}

\begin{table}%
\caption{Simulation Configuration}
\label{tab:one}
\begin{minipage}{\columnwidth}
\begin{center}
\begin{tabular}{llll}
  \toprule
		App.&Tasks&SLOC&Description\\
		\hline
        \emph{ar} &10 &428 & activity recognition using a KNN\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
        \emph{bc} &10 &371 & several bitcount algorithms\\
        % \textbf{bc} &10 &371 & Counts bits in bytes sequences
% using several algorithms and check for correctness\\
        \emph{cuckoo} &14 &426 & Cuckoo Filter with pseudo-random values\\
        \emph{dijkstra} &5 &198 & Dijkstra shortest path algorithm \\
        \emph{fft} &8 &449 & Fast Fourier Transform\\ % on three pre-generated 128 sample vectors\\
		\emph{sort} &4 &167 & selection sort algorithm\\
  \bottomrule
\end{tabular}
\end{center}
\end{minipage}
\caption{Characteristics of benchmarks used for evaluation.}
\label{table:benchmark_table}
\end{table}%

%\begin{table}
%	\centering
%	\footnotesize
%	\begin{tabular}{| r|r|r | p{0.54\columnwidth} |}
%		\hline
%		App.&Tasks&SLOC&Description\\
%		\hline\hline
%        \emph{ar} &10 &428 & activity recognition using a KNN\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
%		\hline
%        \emph{bc} &10 &371 & several bitcount algorithms\\
%        % \textbf{bc} &10 &371 & Counts bits in bytes sequences
%% using several algorithms and check for correctness\\
%		\hline
%        \emph{cuckoo} &14 &426 & Cuckoo Filter with pseudo-random values\\
%		\hline
%        \emph{dijkstra} &5 &198 & Dijkstra shortest path algorithm \\
%		\hline
%        \emph{fft} &8 &449 & Fast Fourier Transform\\ % on three pre-generated 128 sample vectors\\
%		\hline
%		\emph{sort} &4 &167 & selection sort algorithm\\
%		\hline
%	\end{tabular}
%\caption{Characteristics of benchmarks used for evaluation.}
%\label{table:benchmark_table}
%\end{table}

### 2019-06-22 paper/TOSN/intro.tex
TOSN coala version
minor revision TOSN

Upon power failures, a device loses the volatile state in its registers, stack, SRAM, and retains the state of any non-volatile memory, such as FRAM. While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated. These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, leading to an unrecoverable application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based architectural support}~\cite{hicks_isca_2017,idetic,nvp}. Complex hardware architectural changes are expensive to design, verify, and manufacture. New hardware architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software approaches are simpler and applicable to existing devices today. Therefore, this work focuses on software approaches. In particular we address the key limitation of task-based programming and execution model, that is the {\em inflexibility} and \emph{energy-unawareness} of statically decomposing a program into tasks.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/intro-figure-v5.pdf}
    \caption{Task coalescing reduces commit overhead (left), while task splitting enables termination of large tasks (right).
    S: static legacy task-based runtime, C: \sys (this work).}
    % \caption{Task coalescing reduces time and energy overhead by performing fewer commits (denoted as \texttt{C}) than the individual tasks (denoted as \texttt{Tx}) commits (left). On the other hand, task splitting reduces wasted computation and enables termination of big tasks, requiring more energy than the energy buffer can hold (right).}
    \label{fig:coalesce}
\end{figure}

@@

\item[C2] \emph{Dynamic Memory Consistency Handling:} merging static tasks on the fly rises  the need for dynamic memory consistency handling. This leads to the second challenge: how to dynamically detect inter-coalesced-task data dependencies and ensure efficient protection against power interrupts? \sys addresses this challenge by relying on its novel \emph{Virtual Memory Manager} (VMM). The VMM performs real-time dependency tracking to enable protection on a task transition. Individual variables tracking, however, slows down a system dramatically. Therefore, the VMM keeps memory consistent through \emph{privatizing pages} and optimizes optimize bulk shared-data transfer through Direct Memory Access (DMA).

>>>>

Upon power failures, a device loses the volatile state in its registers, stack, SRAM, and retains the state of any non-volatile memory, such as FRAM. While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated. These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, leading to an unrecoverable application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based architectural support}~\cite{hicks_isca_2017,idetic,nvp}. Complex hardware architectural changes are expensive to design, verify, and manufacture. New hardware architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software approaches are simpler and applicable to existing devices today. Therefore, this work focuses on software approaches. In particular we address the key limitation of task-based programming and execution model, that is the {\em inflexibility} and \emph{energy-unawareness} of statically decomposing a program into tasks.
%
\begin{figure}
    \centering
    \includegraphics[width=0.8\columnwidth]{figures/intro-figure-v5.pdf}
    \caption{Task coalescing reduces commit overhead (left), while task splitting enables termination of large tasks (right).
    S: static legacy task-based runtime, C: \sys (this work).}
    % \caption{Task coalescing reduces time and energy overhead by performing fewer commits (denoted as \texttt{C}) than the individual tasks (denoted as \texttt{Tx}) commits (left). On the other hand, task splitting reduces wasted computation and enables termination of big tasks, requiring more energy than the energy buffer can hold (right).}
    \label{fig:coalesce}
\end{figure}

@@

\item[C2] \emph{Dynamic Memory Consistency Handling:} merging static tasks on the fly rises  the need for dynamic memory consistency handling. This leads to the second challenge: how to dynamically detect inter-coalesced-task data dependencies and ensure efficient protection against power interrupts? \sys addresses this challenge by relying on its novel \emph{Virtual Memory Manager} (VMM). The VMM performs real-time dependency tracking to enable protection on a task transition. Individual variables tracking, however, slows down a system dramatically. Therefore, the VMM keeps memory consistent through \emph{privatizing pages} and optimizes bulk shared-data transfer through Direct Memory Access (DMA).
